Composable Concurrency Models

### 2016-11-16 slides/slides.tex
Cleaned up presentation intro/background and typeset diagrams

\usepackage[USenglish]{babel}
\usepackage[backend=bibtex,citestyle=authortitle]{biblatex}
\usepackage[scale=2]{ccicons}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{tikz}

% this actually isn't totally hideous:
\usetheme[numbering=counter, progressbar=frametitle, sectionpage=none]{metropolis}

@@

@@

  \begin{frame}
    \begin{figure}
      \begin{tikzpicture}
        \onslide<1->{
          \node at (0,0) {
            \includegraphics[width=300pt]{browser}
          };
        }

        \onslide<2>{
          \path (-2.25,2.75) [draw=none, fill=darkgray, opacity=0.75] ellipse (60pt and 25pt);
          \node [text=white] at (-2.25,2.75) {
            Multiple tabs
          };

          \draw (2.25,2.25) [draw=none, fill=darkgray, opacity=0.75] ellipse (60pt and 25pt);
          \node [text=white] at (2.25,2.25) {
            Suggestions
          };

          \draw (0,0.5) [draw=none, fill=darkgray, opacity=0.75] ellipse (60pt and 25pt);
          \node [text=white] at (0,0.5) {
            Page rendering
          };

          \draw (-3,-1.5) [draw=none, fill=darkgray, opacity=0.75] ellipse (60pt and 25pt);
          \node [text=white] at (-3,-1.5) {
            Event handling
          };

          \draw (2.5,-1.5) [draw=none, fill=darkgray, opacity=0.75] ellipse (60pt and 25pt);
          \node [text=white] at (2.5,-1.5) {
            Background processes
          };
        }

        \onslide<3>{
          \node at (0,-0.62) {
            \includegraphics[width=220pt]{goals}
          };
        }
      \end{tikzpicture}
    \end{figure}
  \end{frame}

@@

  \begin{frame}
    \frametitle{Concurrency}
    \framesubtitle{Intuitive definition}

    \onslide<1>{
      \begin{columns}[T]
        \begin{column}{0.48\textwidth}
          \centering
          sequential
        \end{column}

        \hfill

        \begin{column}{0.48\textwidth}
          \centering
          concurrent
        \end{column}
      \end{columns}
    }

    \onslide<2>{
      \begin{columns}[T]
        \begin{column}{0.48\textwidth}
          \centering
          concurrent
        \end{column}

        \hfill

        \begin{column}{0.48\textwidth}
          \centering
          parallel
        \end{column}
      \end{columns}
    }
  \end{frame}

  \begin{frame}
    \frametitle{Concurrency}
    \framesubtitle{Formal definition}

@@

@@

  \subsection{Consistency}

  \begin{frame}
    \frametitle{Consistency models}

    \begin{itemize}
      \item \textbf{Sequential:} Does the history of the program yield a correct result?
      \item \textbf{Concurrent:} Does every possible history of the program yield a correct result?
    \end{itemize}
  \end{frame}

@@

    \onslide<1->{
      \centering
      \includegraphics{single-thread-read-write}
    }

    \onslide<2->{
      \centering
      \includegraphics{multiple-threads-read-write}
    }
  \end{frame}

@@

    \begin{figure}
      \centering
      \includegraphics{noninstantaneity}
    \end{figure}
  \end{frame}

  \subsection{Correctness}

@@

    \begin{figure}
      \centering
      \includegraphics{linearizability}
    \end{figure}

@@

    \begin{figure}
      \centering
      \includegraphics{serializability}
    \end{figure}

    \begin{itemize}
      \item Serializability guarantees that operations can occur in any order as long as an equivalent serial history exists.
      \item While a serializable set of operations is being executed, it appears to be the only set of operations being executed.
    \end{itemize}
  \end{frame}

@@

    \begin{itemize}
      \item Linearizability is a guarantee of consistency.
      \item Serializability is a guarantee of isolation.
      \item Together, linearizability and serializability guarantee strict serializability---a history is equivalent to some serial execution order and that order corresponds to the execution order in real time.~\footcite{Herlihy1990}
    \end{itemize}
  \end{frame}

@@

  \begin{frame}[standout]
    \centering

    Thanks to Elena Machkasova, K.K. Lamberty, and Matthew Justin for their guidance and suggestions.

    \url{github.com/dstelljes/senior-sem}

    \ccbyncsa{}
  \end{frame}

>>>>

\usepackage{appendixnumberbeamer}
\usepackage[USenglish]{babel}
\usepackage[backend=bibtex,citestyle=authortitle]{biblatex}
\usepackage[scale=2]{ccicons}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{tikz}

% this actually isn't totally hideous:
\usetheme[numbering=counter, progressbar=frametitle, sectionpage=none]{metropolis}
\setbeamertemplate{bibliography item}[text]
\setbeamertemplate{frametitle continuation}{}

@@

\AtEveryBibitem{\clearfield{note}}
\AtEveryBibitem{\clearfield{url}}

@@

  \begin{frame}
    \centering
    \resizebox{\linewidth}{!}{\input{../figures/introduction.tikz}}
  \end{frame}

@@

  \begin{frame}
    \frametitle{Concurrency}

@@

    If $A \rightarrow B$ and $B \rightarrow C$, then $A \rightarrow C$.

@@

  \subsection{Complications}

@@

    \begin{itemize}
      \item \textbf{Sequential program:} Does the order of operations yield a correct result?
      \vfill
      \item \textbf{Concurrent program:} Does \emph{every possible} order of operations yield a correct result?
    \end{itemize}
  \end{frame}

  \begin{frame}
    \frametitle{Complications}
    \centering

    \textbf{Single thread:}
    \vfill
    \resizebox{0.9\linewidth}{!}{\input{../figures/single-thread-read-write.tikz}}
  \end{frame}

  \begin{frame}
    \frametitle{Complications}
    \centering

    \textbf{Multiple threads:}
    \vfill
    \resizebox{0.9\linewidth}{!}{\input{../figures/multiple-threads-read-write.tikz}}
  \end{frame}

@@

    \centering
    \input{../figures/noninstantaneity.tikz}
  \end{frame}

  \subsection{Consistency models}

@@

    \begin{figure}
      \centering
      \input{../figures/linearizability.tikz}
    \end{figure}

@@

    \begin{figure}
      \centering
      \input{../figures/serializability.tikz}
    \end{figure}

    \begin{itemize}
      \item Serializability guarantees that operations can occur in any order as long as an equivalent sequential ordering exists.
      \item While a serializable set of operations is being executed, it appears to be the only set of operations being executed.
    \end{itemize}
  \end{frame}

@@

    \begin{itemize}
      \item Linearizability is a guarantee of consistency.
      \item Serializability is a guarantee of isolation.
      \item Together, linearizability and serializability guarantee strict serializability---an ordering of operations is equivalent to some sequential ordering and that ordering corresponds to the execution ordering in real time.~\footcite{Herlihy1990}
    \end{itemize}
  \end{frame}

@@

  \begin{frame}[standout]
    \centering
    Thanks to Elena Machkasova, K.K. Lamberty, and Matthew Justin for their guidance and suggestions.
    \vfill
    \href{https://github.com/dstelljes/senior-sem}{github.com/dstelljes/senior-sem}
    \vfill
    \ccbyncsa{}
  \end{frame}

  \appendix{}

  \begin{frame}[allowframebreaks]
    \frametitle{References}

    \printbibliography{}
  \end{frame}
